### 2010-12-03 pieces/z-n-transform/z-n-transform.tex

\documentclass{article}
\usepackage{joyce-math}

\begin{document}
\title{Hello World!}
\author{Joyce Tipping}
\date{}
\maketitle

Let $\Z$ be the standard normal distribution. Show that $\X = \mu + \Z\sigma$ has a normal
distribution with mean $\mu$ and standard deviation $\sigma$.

\begin{proof}
\begin{align*}
M_{\X}(t) & = e^{\mu t} \cdot M_{\Z}(\sigma t) \\
& = e^{\mu t} \cdot e^{\frac{(\sigma t)^2}{2}} \\
& = e^{\mu t + \frac{1}{2}\sigma^2t^2} \\
& = M_{\N}(t)
\end{align*}
\end{proof}
\end{document}