### 2010-12-10 pieces/mabs/mabs.tex
Added skew-normal restrictions graphs.

\documentclass{article}
\usepackage{joyce-math}

\begin{document}
\section{Demonstrating Improved Accuracy}

\subsection{Maximal Absolute Error}

Define the maximal absolute error of the approximation to be

\begin{equation}
  MABS(n, p) \eq \max_{k \in \{0, 1,...,n\}} \left| F_{B(n,p)} (k) -  F_{appr(n,p)}(k + 0.5) \right|
\end{equation}

where $F_{B(n,p)}$ is the cdf of the binomial and $F_{appr(n,p)}$ is the cdf of
either the normal or skew-normal approximation; the 0.5 is a continuity
correction.

\end{document}