### 2010-12-13 pieces/final-paper/output.tex

\documentclass{article}
\usepackage{jcapstone}
\begin{document}

\title{The Skew-Normal Approximation of the Binomial Distribution}
\author{Joyce Tipping}
\date{2010}
\maketitle

\section{Introduction}

One of the most basic distributions in statistics is the binomial, $X \sim
Bin(n,p)$, with pdf

\begin{equation*}
  f_X(x) = \binom{n}{x} p^x q^{n-x}
\end{equation*}

Calculating the binomial cdf, $F_X(x) = P(X \leq x) = \sum_{k=1}^x f_X(k)$, by
hand is manageable for small $n$ but quickly becomes cumbersome as $n$ grows
even mediumly large. A common strategy is to use the normal distribution as an
approximation:

\begin{equation}
  F_X(x) \approx \Phi \left( \frac{k + 0.5 - \mu}{\sigma} \right)
\end{equation}

where $\Phi$ is the standard normal cdf and $\mu = np$ and $\sigma =
\sqrt{np(1-p)}$.

This approximation works well when either $n$ is very large (invoking the
Central Limit Theorem) or when $p$ is close to 0.5 (making $X$ roughly
symmetric). These rules of thumb are captured in a requirement, often stated as

\begin{equation*}
  np(1-p) > 9
\end{equation*}

or

\begin{align*}
      np &> 5 \quad \textnormal{for} \quad  0 < p \leq 0.5, \\
  n(1-p) &> 5 \quad \textnormal{for} \quad  0.5 < p < 1
\end{align*}

However, when $n$ is not large and $p$ is close to 0 or 1, the binomial
distribution is skewed and even if the above requrements are met, the
approximation can be inaccurate; sometimes, as demonstrated by \citet{mabs},
rather substantially. In these cases, the skew-normal approximation can provide
an alternative -- and considerably more accurate -- method of approximation.

\section{The Skew-Normal}

The skew-normal distribution is similar to the normal but with an added
parameter for skew. In this section, we'll introduce

\subsection{Basics}

Let $Y$ be a skew-normal distribution, with location parameter $\mu \in \R$,
scale parameter $\sigma > 0$, and shape parameter $\lambda \in \R$, denoted
$SN(\mu, \sigma, \lambda)$. Then $Y$ has pdf

\begin{equation} \label{eq:sn-pdf}
  f_Y(x) = \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot \Phi \left( \frac{\lambda(x-\mu)}{\sigma} \right)
\end{equation}

where $\phi$ is the standard normal pdf and $\Phi$ is the standard normal cdf.
Some other basic properties of $Y$, given by \citet{pewsey}, are

\begin{align}
  E(Y) &= \mu + b \delta \sigma \nonumber \\
  E(Y^2) &= \mu^2 + 2b \delta \mu \sigma + \sigma^2 \label{eq:sn-basic-properties} \\
  Var(Y) &= \sigma^2 (1 - b^2 \delta^2) \nonumber \\
  E(Y^3) &= \mu^3 + 3 b \delta \mu^2 \sigma + 3 \mu \sigma^2 + 3 b \delta \sigma^3 - b \delta^3 \sigma^3 \nonumber
\end{align}

where $b = \sqrt{\frac{2}{\pi}}$ and $\delta = \frac{\lambda}{\sqrt{1 +
\lambda^2}}$.

The $SN(0,1,\lambda)$ distribution is called the standard skew-normal, and has
pdf

\begin{equation} \label{eq:standard-sn-pdf}
  f_z(x) = 2 \cdot \phi(x) \cdot \Phi (-\lambda x)
\end{equation}

Like the normal and standard normal, you can arrive at the standard skew normal
by applying the transformation $\frac{Y - \mu}{\sigma}$.

A natural question to ask is how the skew-normal relates to the normal.
Fortunately, the connection is very intuitive: When $\lambda = 0$, Equation
\eqref{eq:sn-pdf} reverts to the normal pdf:

\begin{align*}
  f_Y(x|\lambda=0) &= \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot \Phi(0) \\
  &= \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot 0.5 \\
  &= \frac1\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \\
  &= \frac1\sigma \cdot \frac{1}{\sqrt{2\pi}} \;\cdot\; \exp \left( -\frac{(x-\mu)^2}{2\sigma^2} \right) \\
  &= \frac{1}{\sqrt{2\pi\sigma^2}} \;\cdot\; \exp \left( -\frac{(x-\mu)^2}{2\sigma^2} \right)
\end{align*}

\subsection{Four Properties}
\label{subsec:four-properties}

\citet{main} gives four handy properties of the skew-normal distribution:

\begin{property} \label{prop:1}
  If $Z \sim SN(0, 1, \lambda)$, then $(-Z) \sim SN(0, 1, -\lambda)$.
\end{property}

\begin{proof}
  The standard normal pdf is an even function: $\phi(-x) =
  \frac{1}{\sqrt{2\pi}}\;e^{-(-x)^2/2} = \frac{1}{\sqrt{2\pi}}\;e^{-x^2/2} =
  \phi(x)$. But the standard normal cdf, \thinspace $\Phi(x) =
  \int_{-\infty}^\infty \phi(x)$, \thinspace is not, being 0 near $-\infty$ and
  1 near $\infty$. Thus,

  \begin{align*}
    f_{(-Z)}(x) &= f_Z(-x) \\
    & = 2 \cdot \phi(-x) \cdot \Phi (-\lambda x) \\
    & = 2 \cdot \phi(x) \cdot \Phi (-\lambda x)
  \end{align*}

  which is the pdf of $SN(0, 1, -\lambda)$.
\end{proof}

\begin{property} \label{prop:2}
  As $\lambda \to \pm \infty$, \thinspace $SN(0,1,\lambda)$ tends to the half normal distribution, $\pm |N(0,1)|$.
\end{property}

To prove our theorem, it is helpful to formally define the half normal distribution:

\begin{helper-lem} \label{lem:p2-half-normal}
  Let $X \sim N(0, \sigma^2)$. Then the distribution of $|X|$ is a half-normal
  random variable with parameter $\sigma$ and

  \begin{equation*}
    f_{|X|}(x) =
    \begin{dcases*}
      2 \cdot f_X(x) & when $0 < x < \infty$ \\
      0 & everywhere else
    \end{dcases*}
  \end{equation*}
\end{helper-lem}

\begin{proof}
  Let $X \sim N(0, \sigma^2)$, defined over $A = (-\infty, \infty)$. Define

  \begin{equation*}
    Y = |X| =
    \begin{dcases*}
      -x & if $x < 0$ \\
      0 & if $x = 0$ \\
      x & if $x > 0$ \\
    \end{dcases*}
  \end{equation*}

  $Y$ is not one-to-one over $A$. However, we can partition $A$ into disjoint
  subsets $A_1 = (-\infty, 0)$, $A_2 = (0, \infty)$, and $A_3 = \{0\}$ such
  that $A = A_1 \cup A_2 \cup A_3$ and $Y$ is one-to-one over each $A_i$. We
  can then transform each piece separately using Theorem 6.3.2 from
  \citet{textbook}:

  On $A_1$: $y = -x \lra x = -y$ and $\J = \left| \frac{dx}{dy} \right| = |-1|
  = 1$, yielding

  \begin{align}
    f_Y(y) &= f_X(x) \cdot \J \nonumber \\
    &= f_X(-y) \cdot 1 \nonumber \\
    &= \frac{1}{\sqrt{2\pi}\sigma} \; e^{-\frac{(-y)^2}{2\sigma}} \nonumber \\
    &= \frac{1}{\sqrt{2\pi}\sigma} \; e^{-\frac{y^2}{2\sigma}} \nonumber \\
    &= f_X(y) \label{eq:p2-transform-a1}
  \end{align}

  over the domain $A_1: -\infty < x < 0 \lra -\infty < -y < 0 \lra 0 < y < \infty :B_1$.

  Similarly, on $A_2$: $y = x \lra x = y$ and $\J = \left| \frac{dx}{dy}
  \right| = |1| = 1$, yielding

  \begin{align}
    f_Y(y) &= f_X(x) \cdot \J \nonumber \\
    &= f_X(y) \cdot 1 \nonumber \\
    &= f_X(y) \label{eq:p2-transform-a2}
  \end{align}

  over the domain $A_2: 0 < x < \infty \lra 0 < y < \infty :B_2$.

  On $A_3$, we have $x = 0, y = 0$ and $\J = \left| \frac{dx}{dy} \right| = |0|
  = 0$, yielding $f_Y(y) = f_X(x) \cdot \J = f_X(x) \cdot 0 = 0$.

  Now, by Theorem 6.3.10 from \citet{textbook}, we achieve our result by simply
  summing \eqref{eq:p2-transform-a1} and \eqref{eq:p2-transform-a2}.

  \begin{align}
    f_Y(y) &= \{ f_Y(y) \textrm{ over } A_1 \} + \{ f_Y(y) \textrm{ over } A_2 \} \nonumber \\
    &= f_X(y) + f_X(y) \nonumber \\
    &= 2 \cdot f_X(y) \label{eq:p2-transform-result}
  \end{align}

  over $B = B_1 \cup B_2 = (0, \infty)$, and 0 otherwise.
\end{proof}

With Lemma \ref{lem:p2-half-normal}, we can easily show our property:

\begin{proof}[Proof of Property 2]
  Let $Z \sim SN(0,1,\lambda)$. Recall that $f_z(x) = 2 \cdot \phi(x) \cdot
  \Phi(\lambda x)$.

  Consider $\lim_{\lambda \to \infty} f_X(x)$. When $x$ is negative, $\lambda x
  \to -\infty$ and thus $\Phi(\lambda x) \to 0$. When $x$ is positive, however,
  $\lambda x \to \infty$ and $\Phi(\lambda x) \to 1$. Thus

  \begin{equation}
    \label{eq:p2-positive-half-normal}
    \lim_{\lambda \to \infty} 2 \cdot \phi(x) \cdot \Phi(\lambda x) \eq
    \begin{dcases*}
      0 & when $x \leq 0$ \\
      2 \cdot \phi(x) & when $x > 0$
    \end{dcases*}
    \eq |N(0,1)|
  \end{equation}

  In $\lim_{\lambda \to -\infty} f_X(x)$, the signs are reversed. When $x$ is
  negative, $\lambda x \to \infty$ and $\Phi(\lambda x) \to 1$. When $x$ is
  positive, $\lambda x \to -\infty$ and $\Phi(\lambda x) \to 0$. Thus,

  \begin{equation}
    \label{eq:p2-negative-half-normal}
    \lim_{\lambda \to -\infty} 2 \cdot \phi(x) \cdot \Phi(\lambda x) \eq
    \begin{dcases*}
      2 \cdot \phi(x) & when $x < 0$ \\
      0 & when $x \geq 0$
    \end{dcases*}
    \eq -|N(0,1)|
  \end{equation}
\end{proof}

\begin{property} \label{prop:3}
  If $Z \sim SN(0, 1, \lambda)$, then $Z^2 \sim \chi^2_1$ (chi-square with 1 degree of freedom).
\end{property}

\begin{proof}
  To prove our result, we make use of Lemma 1 in \citet{azzalini}:

  \begin{helper-lem}
    If $f_0$ is a one-dimensional probability density function symmetric about
    0, and $G$ is a one-dimensional distribution function such that $G'$ exists
    and is a density symmetric about 0, then

    \begin{equation}
      \label{eq:azzalini-perturbation-invariance}
      f(z) = 2 \cdot f_0(z) \cdot G\{w(z)\} \quad (-\infty < z < \infty)
    \end{equation}

    is a density function for any odd function $w(\cdot)$.
  \end{helper-lem}

  Notice that $\phi(x)$ is a one-dimensional probability density function
  symmetric about 0, and $\Phi(x)$ is a one-dimensional distribution function
  such that $\Phi'$ exists and is a density symmetric about 0. Furthermore,
  $\lambda x$ is an odd function. Thus, $f_z(z) = 2 \cdot \phi(z) \cdot
  \Phi(\lambda z)$ conforms to equation
  \eqref{eq:azzalini-perturbation-invariance}. With that in mind, the corollary
  to this lemma provides a very useful result:

  \begin{helper-cor}[Perturbation Invariance]
    If $Y \sim f_0$ and $Z \sim f$, then $|Y| \overset{d}{=} |Z|$, where the
    notation $\overset{d}{=}$ denotes equality in distribution.    
  \end{helper-cor}

  Thus, we can treat $\phi$ and $Z$ as being equal in distribution. We will now
  show that $\phi^2 \sim \chi^2_1$:

  \begin{align*}
    M_{\phi^2}(t) &= E[e^{tx^2}] \\
    &= \int_{-\infty}^\infty \; e^{tx^2} \left[ \frac{1}{\sqrt{2\pi}} e^{-x^2/2} \right] dx \\
    &= \int_{-\infty}^\infty \; \frac{1}{\sqrt{2\pi}} \; e^{tx^2 - x^2/2} \; dx \\
    &= \int_{-\infty}^\infty \; \frac{1}{\sqrt{2\pi}} \; e^{-\frac{x^2}{2}(1 - 2t)} \; dx \\
    &= \int_{-\infty}^\infty \; \frac{1}{\sqrt{2\pi}} \; e^{-\frac{1}{2}(\sqrt{1-2t} \; x)^2} \; dx \\
    \intertext{Let $u = (\sqrt{1-2t}) \, x$; then we have $du = \sqrt{1-2t}$, \enspace $dx = \frac{du}{\sqrt{1-2t}}$, \enspace and our limits become $x \to -\infty \Ra u \to -\infty$
      and $x \to \infty \Ra u \to \infty$.}
    &= \int_{-\infty}^\infty \; \frac{1}{\sqrt{2\pi}} \; e^{-u^2/2} \; \left( \frac{1}{\sqrt{1-2t}} \right) du \\
    &= \frac{1}{\sqrt{1-2t}} \; \underbrace{\left( \int_{-\infty}^\infty \; \frac{1}{\sqrt{2\pi}} \; e^{-u^2/2} \; du \right)}_{\mathclap{\textnormal{$\phi(u)$ integrated over
      $(-\infty,\infty)$ = 1}}} \label{phi-pdf} \\
    &= \frac{1}{\sqrt{1-2t}}
  \end{align*}

  which is the MGF of the $\chi^2_1$. Since $Z$ is equal in distribution to
  $\phi$, we can also conclude that $Z^2 \sim \chi^2_1$. \end{proof}

\begin{property} \label{prop:4}
  The MGF of $SN(0,1,\lambda)$ is

  \begin{equation} \label{eq:p4-sn-mgf}
    M(t|\lambda) = 2 \cdot \Phi (\delta t) \cdot e^{t^2/2}
  \end{equation}

  where $\delta = \frac{\lambda}{\sqrt{1 + \lambda^2}}$ and $t \in (-\infty, \infty)$.
\end{property}

\begin{proof}
  According to Equation 5 in \citet{azzalini}, the MGF of $SN(\mu, \sigma^2,
  \lambda)$ is

  \begin{equation*}
    M(t) \eq E\{e^{tY}\} \eq 2 \cdot \exp \left( \mu t + \frac{\sigma^2 t^2}{2} \right) \cdot \Phi(\delta \sigma t)
  \end{equation*}

  where $\delta = \frac{\lambda}{1 + \lambda^2} \in (-1, 1)$. It follows that
  the MGF of the $SN(0, 1, \lambda)$ is

  \begin{equation*}
    2 \cdot \exp \left( 0 \cdot t + \frac{1 \cdot t^2}{2} \right) \cdot \Phi(\delta \cdot 1 \cdot t) \eq 2 \cdot e^{t^2/2} \cdot \Phi(\delta t)
  \end{equation*}
\end{proof}

\section{Developing an Approximation}

Next, we develop our skew-normal approximation of the binomial. Let $B \sim
Bin(n, p)$ and $Y \sim SN(\mu, \sigma^2, \lambda)$. We will find estimates for
$\mu$, $\sigma$, and $\lambda$ by comparing their first, second, and third
moments about the mean.

\subsection{The Moments of the Binomial}

We begin by examining the moments of the binomial. The first two, the mean and
variance, are simply

\begin{equation*}
  E(B) = np, \quad Var(B) = np(1-p)
\end{equation*}

Having these, we can easily find

\begin{equation*}
  E(B^2) = Var(B) + [E(B)]^2 = np(1-p) + n^2p^2 = np - np^2 + n^2p^2
\end{equation*}

which we will need for the third moment. We will also need $E(B^3)$, which we
will get via the third factorial moment:

\begin{align*}
  E[B(B-1)(B-2)] &= \sum_{x=0}^n x (x-1) (x-2) \cdot \left\{ \binom{n}{x} p^x q^{n-x} \right\} \\
  &= \sum_{x=3}^n x(x-1)(x-2) \cdot \frac{n!}{x!\;(n-x)!} \; p^x q^{n-x} \\
  &= \sum_{x=3}^n \frac{n!}{(x-3)!\;(n-x)!} \; p^x q^{n-x} \\
  &= \sum_{x=3}^n n(n-1)(n-2) p^3 \cdot \frac{(n-3)!}{(x-3)!\;(n-x)!} \; p^{x-3}q^{n-x} \\
  \intertext{Let $y=x-3$. Then $x=y+3$, and $x=3 \rightarrow y=0$ and $x=n \rightarrow y=n-3$.}
  &= n(n-1)(n-2)p^3 \cdot \sum_{y=0}^{n-3} \frac{(n-3)!}{y!\;(n-(y+3))!} \; p^y q^{n-(y+3)} \\
  &= n(n-1)(n-2)p^3 \cdot \underbrace {\sum_{y=0}^{n-3} \frac{(n-3)!}{y!\;((n-3)-y)!} \; p^y q^{(n-3)-y}}_{\mathclap{\textnormal{[pdf of $Bin(n-3,p)$ summed from 0 to $n-3$] = 1}}} \\
  &= n(n-1)(n-2)p^3 \\
  &= n^3p^3 - 3n^2p^3 + 2np^3 \\
  \intertext{Further expanding the left side and solving for $E(B^3)$,}
  E \left[ B^3 - 3B^2 + 2B \right] &= n^3p^3 - 3n^2p^3 + 2np^3 \\
  E(B^3) - 3E(B^2) + 2E(B) &= \\
  E(B^3) - 3(np - np^2 + n^2p^2) + 2np &= \\
  \Rightarrow \qquad E(B^3) &= n^3p^3 - 3n^2p^3 + 2np^3 + 3np - 3np^2 + 3n^2p^2 - 2np \\
  &= n^3p^3 - 3n^2p^3 + 2np^3 - 3np^2 + 3n^2p^2 + np
\end{align*}

With these results (and a bit of elbow grease), we can obtain the third moment
without too much trouble:

\begin{align*}
  E \left( [B - E(B)]^3 \right) &= E \left( B^3 - 3B^2 E(B) + 3B [E(B)]^2 - [E(B)]^3 \right) \\
  &= E(B^3) - 3 E(B^2) E(B) + 3 E(B) [E(B)]^2 - [E(B)]^3 \\
  &= E(B^3) - 3 E(B^2) E(B) + 2 [E(B)]^3 \\
  &= (n^3p^3 - 3n^2p^3 + 2np^3 - 3np^2 + 3n^2p^2 + np) - 3np(np - np^2 + n^2p^2) + 2n^3p^3 \\
  &= \cancel{n^3p^3} - \cancel{3n^2p^3} + 2np^3 - 3np^2 + \cancel{3n^2p^2} + np - \cancel{3n^2p^2} + \cancel{3n^2p^3} - \cancel{3n^3p^3} + \cancel{2n^3p^3} \\
  &= 2np^3 - 3np^2 + np \\
  &= np(p-1)(2p-1)
\end{align*}

Our hard-earned results, restated for convenience:

\begin{align}
  E(B) &= np \nonumber \\
  E([B - E(B)]^2) &= np(1-p) \\
  E([B - E(B)]^3) &= np(p-1)(2p-1) \nonumber
\end{align}

\subsection{The Moments of the Skew Normal}

Now we'll take a look at the moments of the skew normal. Equation
\eqref{eq:sn-basic-properties} takes care of the mean and variance; again the
third moment is a little more complicated:

\begin{align*}
  E([Y - E(Y)]^3) &= E(Y^3) - 3E(Y^2)E(Y) + 2[E(Y)]^3 \\
  &= (\mu^3 + 3 b \delta \mu^2 \sigma + 3 \mu \sigma^2 + 3 b \delta \sigma^3 - b \delta^3 \sigma^3) - 3 (\mu^2 + 2b \delta \mu \sigma + \sigma^2) (\mu + b \delta \sigma) \\
  & \quad + 2\;(\mu + b \delta \sigma)^3 \\
  &= \cancel{\mu^3} + \cancel{3 b \delta \mu^2 \sigma} + \cancel{3 \mu \sigma^2} + \cancel{3 b \delta \sigma^3} - b \delta^3 \sigma^3 - \cancel{3 \mu^3} - \cancel{9 b \delta \mu^2 \sigma} -
    \cancel{6 b^2 \delta^2 \mu \sigma^2} - \cancel{3 \mu \sigma^2} \\
  & \quad - \cancel{3 b \delta \sigma^3} + \cancel{2 \mu^3} + \cancel{6 b \delta \mu^2 \sigma} + \cancel{6 b^2 \delta^2 \mu \sigma^2} + 2 b^3 \delta^3 \sigma^3 \\
  &= 2 b^3 \delta^3 \sigma^3 - b \delta^3 \sigma^3 \\
  &= b \delta^3 \sigma^3 (2b^2 - 1)
\end{align*}

We restate our results:

\begin{alignat}{4}
  E(Y) &= \mu + b \delta \sigma \;&=&\; \mu + \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}} \nonumber \\
  E([Y - E(Y)]^2) &= \sigma^2 (1 - b^2 \delta^2) \;&=&\; \sigma^2 \left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2} \right) \\
  E([Y - E(Y)]^3) &= b \delta^3 \sigma^3 (2b^2 - 1) \;&=&\; \sigma^3 \sqrt{\frac{2}{\pi}} \left( \frac{\lambda}{\sqrt{1 + \lambda^2}} \right)^3 \left( \frac{4}{\pi} - 1 \right) \nonumber
\end{alignat}

\subsection{Solving for $\mu$, $\sigma$, $\lambda$}

Now we set the two sets of moments equal to each other and, taking $n$ and $p$
as constants, solve for $\mu$, $\sigma$ and $\lambda$.

\begin{subequations}
\begin{align}
  np &= \mu + \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}} \label{eq:first-moment-set} \\
  np(1-p) &= \sigma^2 \left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2} \right) \label{eq:second-moment-set} \\
  np(p-1)(2p-1) &= \sigma^3 \sqrt{\frac{2}{\pi}} \left( \frac{\lambda}{\sqrt{1 + \lambda^2}} \right)^3 \left( \frac{4}{\pi} - 1 \right) \label{eq:third-moment-set}
\end{align}
\end{subequations}

To get $\lambda$, we divide the cube of \eqref{eq:second-moment-set} by the
square of \eqref{eq:third-moment-set}:

\begin{align}
  \frac{\sigma^6 \left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2} \right)^3}{\sigma^6 \cdot \frac{2}{\pi} \left( \frac{\lambda}{\sqrt{1 + \lambda^2}} \right)^6 \left(
    \frac{4}{\pi} - 1 \right)^2} &= \frac{n^3p^3(1-p)^3}{n^2p^2(p-1)^2(2p-1)^2} \nonumber \\
  \Rightarrow \quad \frac{\left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1+\lambda^2} \right)^3}{\frac{2}{\pi} \left( \frac{\lambda^2}{1+\lambda^2} \right)^3 \left( \frac{4}{\pi} - 1
    \right)^2} &= \frac{np(1-p)}{(1-2p)^2} \label{eq:solving-for-lambda}
\end{align}

The above equation \eqref{eq:solving-for-lambda} is a rational expression in
$\lambda^2$ that can be solved with either a considerable amount of manual
labor or, more efficiently, with a computer algebra system. Once we have
$\lambda^2$, then $\lambda$ is simply either the positive or negative square
root, as determined by the sign of $(1-2p)$. This can be explained with a
little assistance from Property \ref{prop:3}: When $p \to 0$, the binomial
skews left and converges toward the positive half normal, which by
\eqref{eq:p2-positive-half-normal} corresponds to a positive $\lambda$. When $p
\to 1$, the binomial skews right and converges toward the negative half normal,
which by \eqref{eq:p2-negative-half-normal} corresponds to a negative
$\lambda$. When $p = 0.5$, the binomial is symmetric and $\lambda$ is 0,
eliminating the need for a sign. Thus:

\begin{equation}
  \label{eq:lambda-solved}
  \lambda = \textnormal{\{sign of $(1-2p)$\}} \sqrt{\lambda^2}
\end{equation}

%\textbf{\textit{Note to Dr. Guffey:} I am trying to find an algebraic reason
%for the sign, but I'm not seeing it in equation \eqref{eq:solving-for-lambda}
%nor its algebraic expansion, which I put in in appendix
%\ref{sec:solving-for-lambda} as it ended up providing little insight into this
%conclusion.}

Having secured $\lambda$, we can find $\sigma$ using
\eqref{eq:second-moment-set}:

\begin{equation}
  \label{eq:sigma-solved}
  np(1-p) = \sigma^2 \left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2} \right) \quad\Rightarrow\quad
  \sigma = \sqrt{\frac{np(1-p)}{1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2}}}
\end{equation}

Similarly, with both $\lambda$ and $\sigma$, a simple rearrangement of
\eqref{eq:first-moment-set} yields $\mu$:

\begin{equation}
  \label{eq:mu-solved}
  np = \mu + \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}} \quad\Rightarrow\quad
  \mu = np - \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}}
\end{equation}

When $p = 0.5$, we would expect the binomial to be perfectly symmetrical and
the mean therefore to be $n/2$. From \eqref{eq:mu-solved}, this implies that
$\sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}} =
0 \Ra$ either $\sigma = 0$ or $\lambda = 0$. Since the former is impossible, we
must conclude the latter, which brings us back to the normal distribution.

%\textbf{\textit{Another note to Dr. Guffey:} The author mentions here that when
%$p=0.5$, it forces $\lambda$ to be 0, bringing us back to the regular normal
%approximation. This seems like it ought to be a simple fact, but I can't figure
%out how he's getting it from these equations. He refers to his equation (c),
%which is my equation \eqref{eq:mu-solved}. Substituting $p=0.5$ into
%\eqref{eq:mu-solved}, however, doesn't get me anywhere; neither does
%substituting $p=0.5$ into \eqref{eq:sigma-solved} and then substituting that
%into \eqref{eq:mu-solved}. And as far as I can tell, equation
%\eqref{eq:lambda-solved} is for all practical purposes useless. Alas.}

\subsection{Restrictions}

To obtain an estimate for $\lambda$, we must put a few restrictions on $n$ and
$p$.

If we let $u = \frac{\lambda^2}{1+\lambda^2}$ and $v = 1/u$, we can rewrite the
left hand side of \eqref{eq:solving-for-lambda} as

\begin{gather}
  \left. \left( 1 - \frac{2}{\pi} u \right)^3 \middle/ \frac{2}{\pi} u^3 \left( \frac{4}{\pi} - 1 \right)^2 \right. \nonumber \\
  \left( 1 - \frac{2}{\pi} u \right)^3 \cdot v^3 \cdot \frac{\pi}{2} \cdot \left( \frac{\pi}{4-\pi} \right)^2 \nonumber \\
  \left[ v \left( 1 - \frac{2}{\pi} u \right) \right]^3 \left( \frac{\pi^3}{2(4-\pi)^2} \right) \nonumber \\
  \left( v - \frac{2}{\pi} \right)^3 \left( \frac{\pi^3}{2(4-\pi)^2} \right) \label{eq:lhs-solving-for-lambda}
\end{gather}

we can see that it is increasing in $v$, which is always $\geq 1$. Therefore:

\begin{equation}
  \min_{v} \textnormal{\{Eq. \ref{eq:lhs-solving-for-lambda}\}} = \textnormal{\{Eq. \ref{eq:lhs-solving-for-lambda}\}} |_{v=1} = 
  \left( 1 - \frac{2}{\pi} \right)^3 \left( \frac{\pi^3}{2(4-\pi)^2} \right) = 1.009524 \approx 1
\end{equation}

This means that the right hand side of \eqref{eq:solving-for-lambda}, which is
supposed to be equal to the left hand side of \eqref{eq:solving-for-lambda},
can't ever be less than 1. Unfortunately, it sometimes is; in particular,
$\frac{np(1-p)}{(1-2p)^2} \to 0$ both when $p \to 0$ and $p \to 1$. So if we
want a solution, we must restrict $n$ and $p$ such that

\begin{align}
  \textnormal{\{right hand side of \eqref{eq:solving-for-lambda}\}} &\geq \textnormal{\{min of left hand side of \eqref{eq:solving-for-lambda}\}} \nonumber \\
  \frac{np(1-p)}{(1-2p)^2} &\geq 1 \nonumber \\
  np(1-p) &\geq (1-2p)^2 \label{eq:solving-the-restriction}
\end{align}

From here, given a constant $p$, finding $n$ is very simple:

\begin{equation}
  n \geq \frac{(1-2p)^2}{p(1-p)} \label{eq: n for a given p}
\end{equation}

Figure \ref{fig:sn-restriction-least-n} shows the least sample size required to
estimate $\lambda$, given a fixed $p$. As expected, the least $n$ is quite
large when $p$ is small and $\to 0$ as $p \to 0.5$. For example, when $p =
0.01$, $n \geq 98$; but at $p = 0.2$, $n$ must only be $\geq 3$, a trivial
requirement to meet.

\begin{figure}
  \centering
  \subfloat[Least possible $n$, given a fixed $p$]{\label{fig:sn-restriction-least-n}\includegraphics[width=0.45\textwidth]{../graphs/images/restriction-least-n.png}}
  \quad
  \subfloat[Range of possible $p$, given a fixed $n$]{\label{fig:sn-restriction-p-range}\includegraphics[width=0.45\textwidth]{../graphs/images/restriction-p-range.png}}
  \caption{Restrictions on $n$ and $p$ for estimating $\lambda$}
  \label{fig:sn-restriction}
\end{figure}

It is also possible to fix $n$ and solve for $p$. We return to
\eqref{eq:solving-the-restriction} for further factoring

\begin{align}
  np - np^2 &\geq 1 - 4p + 4p^2 \nonumber \\
  1 - 4p + 4p^2 - np + np^2 &\leq 0 \nonumber \\
  (n+4)p^2 - (n+4)p + 1 &\leq 0 \label{eq: solving for p}
\end{align}

and apply the quadratic formula with $a = n+4$, $b = -(n+4)$, and $c = 1$:

\begin{gather*}
  \frac{(n+4) \pm \sqrt{(n+4)^2 - 4 \cdot (n+4) \cdot 1}}{2(n+4)} \\
  \frac{(n+4) \pm \sqrt{n^2 + 8n + 16 - 4n - 16}}{2(n+4)} \\
  \frac{(n+4) \pm \sqrt{n^2 + 4n}}{2(n-4)} \\
  \frac{n+4}{2(n+4)} \pm \frac12 \sqrt{\frac{n(n+4)}{(n+4)^2}} \\
  \frac12 \pm \frac12 \sqrt{\frac{n}{n+4}}
\end{gather*}

Let $r_1 = \frac12 - \frac12 \sqrt{\frac{n}{n+4}}$ and $r_2 = \frac12 + \frac12
\sqrt{\frac{n}{n+4}}$. (Note that $r_1 < r_2$.) Now we can rewrite \eqref{eq:
solving for p} as

\begin{equation*}
  (p - r_1)(p - r_2) \leq 0
\end{equation*}

Examining the left hand side, when $p < r_1$, both terms are negative and so
their product is positive; when $p > r_2$, both terms are positive, again
leading the product to be positive. Therefore, our solution lies where $r_1
\leq p \leq r_2$, or more explicitly:

\begin{equation}
 \frac12 - \frac12 \sqrt{\frac{n}{n+4}} \; \leq \; p \; \leq \; \frac12 + \frac12 \sqrt{\frac{n}{n+4}}
\end{equation}

As shown in figure \ref{fig:sn-restriction-p-range}, this interval grows
quickly as $n$ increases, and for sufficiently large $n$, it becomes almost
$(0, 1)$. For example, when $n=100$, our interval is $(0.00971, 0.99029)$; when
$n=500$, it is (0.00199, 0.99801).

For $p$ so close to 0 or 1 that this solution will not work, our authors
suggest a Poission approximation.

\section{Demonstrating Improved Accuracy}

Having taken the trouble to develop our skew-normal approximation for the
binomial, we now justify our efforts by demonstrating its improved accuracy
over the regular normal approximation.

\subsection{Visual Comparison}

The first, and most obvious, way of judging accuracy is by visual inspection.
Figures \ref{fig:comparison-n25}, \ref{fig:comparison-n50}, and
\ref{fig:comparison-n100} compare the binomial, normal, and skew-normal at
small values of $p$ for $n=25$, $n=50$, and $n=100$, respectively. It is not
hard to see that, especially at very small $n$ and $p$, our skew-normal curve
follows the shape of the binomial much more accurately.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../graphs/images/comparison-n25.png}
  \caption{Binomial, normal, and skew-normal, $n=25$}
  \label{fig:comparison-n25}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../graphs/images/comparison-n50.png}
  \caption{Binomial, normal, and skew-normal, $n=50$}
  \label{fig:comparison-n50}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../graphs/images/comparison-n100.png}
  \caption{Binomial, normal, and skew-normal, $n=100$}
  \label{fig:comparison-n100}
\end{figure}

\subsection{Maximal Absolute Error}

Another more numerical method would be to compare the maximal absolute errors
of our two approximations, defined by \citet{mabs} as

\begin{equation}
  \textnormal{MABS}(n, p) \eq \max_{k \in \{0, 1,...,n\}} \left| F_{B(n,p)} (k) -  F_{\textnormal{appr}(n,p)}(k + 0.5) \right|
\end{equation}

where $F_{B(n,p)}$ is the cdf of the binomial and $F_{\textnormal{appr}(n,p)}$
is the cdf of either the normal or skew-normal approximation; the 0.5 is a
continuity correction.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../graphs/images/mabs-fixed-n.png}
  \caption{MABS as a function of p}
  \label{fig:mabs-fixed-n}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{../graphs/images/mabs-fixed-p.png}
  \caption{MABS as a function of n}
  \label{fig:mabs-fixed-p}
\end{figure}

Figure \ref{fig:mabs-fixed-n} shows the MABS as a function of $p$, for $n=25$
and $n=100$. Figure \ref{fig:mabs-fixed-p}, on the other hand, shows the MABS
as a function of $n$, for $p=0.05$ and $p=0.1$. Again, the skew-normal
outperforms the normal considerably in the extreme ranges, with the two
approximations converging as $n \rightarrow \infty$ or $p \rightarrow 0.5$.

\clearpage

\bibliography{/home/joycetipping/projects/capstone/pieces/bibliography/bibliography.bib}
\bibliographystyle{plainnat}
\end{document}