### 2010-12-24 pieces/final-paper/intro.tex
Added punctuation, support sets, made several small tweaks as suggested by Dr. Guffey.

\title{The Skew-Normal Approximation of the Binomial Distribution}
\author{Joyce Tipping}
\date{2010}
\maketitle

\section{Introduction}

One of the most basic distributions in statistics is the binomial, $X \sim
Bin(n,p)$, with pdf

\begin{equation*}
  f_X(x) = \binom{n}{x} p^x q^{n-x}, \quad x = 0, 1, \ldots, n.
\end{equation*}

Calculating the binomial cdf, $F_X(x) = P(X \leq x) = \sum_{k=1}^x f_X(k)$, by
hand is manageable for small $n$ but quickly becomes cumbersome as $n$ grows
even moderately large. A common strategy is to use the normal distribution as
an approximation:

\begin{equation}
  F_X(x) \approx \Phi \left( \frac{x + 0.5 - \mu}{\sigma} \right),
\end{equation}

where $\Phi$ is the standard normal cdf and $\mu = np$ and $\sigma =
\sqrt{np(1-p)}$.

This approximation works best when the binomial is perfectly symmetrical, with
$p=0.5$. However as $p$ travels away from 0.5 in either direction, the binomial
becomes increasingly skewed, and one must either compensate by providing ever
larger values of $n$ or face growing inaccuracy. In these cases, the
skew-normal approximation can provide a better alternative.