### 2010-12-25 project/paper/intro.tex
Proofread new appendix section.

\title{The Skew-Normal Approximation of the Binomial Distribution}
\author{Joyce Tipping}
\date{2010}
\maketitle

\section{Introduction}

One of the most basic distributions in statistics is the binomial, $X \sim
Bin(n,p)$, $n \in \N$, $p \in (0, 1)$ with pdf

\begin{equation*}
  f_X(x) = \binom{n}{x} p^x q^{n-x}, \quad x = 0, 1, \ldots, n,
\end{equation*}

where $q=1-p$. Calculating the binomial cdf, $F_X(x) = P(X \leq x) =
\sum_{k=1}^x f_X(k)$, by hand is manageable for small $n$ but quickly becomes
cumbersome as $n$ grows even moderately large. A common strategy is to use the
normal distribution as an approximation:

\begin{equation}
  F_X(x) \approx \Phi \left( \frac{x + 0.5 - \mu}{\sigma} \right),
\end{equation}

where $\Phi$ is the standard normal cdf and $\mu = np$ and $\sigma =
\sqrt{np(1-p)}$.

This approximation works best when the binomial is perfectly symmetrical, with
$p=0.5$. However as $p$ travels away from 0.5 in either direction, the binomial
becomes increasingly skewed, and one must either provide larger values of $n$
to compensate or face growing inaccuracy. In these cases, the skew-normal
approximation can provide a better alternative.