### 2011-02-13 project/presentation/skew-normal.tex

% The Skew-Normal Distribution
\frame{
  \frametitle{The Skew-Normal Distribution}
  \begin{definition}[Skew-normal]
    Let $Y$ be a skew-normal distribution, with location parameter $\mu \in
    \R$, scale parameter $\sigma > 0$, and shape parameter $\lambda \in \R$.
    Then $Y$ has pdf

    \begin{equation*}
      f(x|\mu, \sigma, \lambda) = \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot \Phi \left( \frac{\lambda(x-\mu)}{\sigma} \right), \quad x \in \R,
    \end{equation*}

    where $\phi$ is the standard normal pdf and $\Phi$ is the standard normal
    cdf.

    \vsep
    We write $Y \sim SN(\mu, \sigma, \lambda)$.
  \end{definition}
}
\frame{
  \frametitle{The Skew-Normal Distribution}
  \begin{lem}
    \label{lem:azzalini-lemma-1}

    If $f_0$ is a one-dimensional probability density function symmetric about
    0, and $G$ is a one-dimensional distribution function such that $G'$ exists
    and is a density symmetric about 0, then

    \begin{equation}
      \label{eq:azzalini-skewed-distribution}
      f(z) = 2 \cdot f_0(z) \cdot G\{w(z)\} \quad (-\infty < z < \infty)
    \end{equation}

    is a density function for any odd function $w(\cdot)$. (Lemma 1,
    \citealp{azzalini})
  \end{lem}
}
\frame{
  \frametitle{The Skew-Normal Distribution}

  \begin{align*}
    E(Y) &= \mu + b \delta \sigma, \\
    E(Y^2) &= \mu^2 + 2b \delta \mu \sigma + \sigma^2, \label{eq:sn-moments-about-origin} \\
    E(Y^3) &= \mu^3 + 3 b \delta \mu^2 \sigma + 3 \mu \sigma^2 + 3 b \delta \sigma^3 - b \delta^3 \sigma^3, \\
    Var(Y) &= \sigma^2 (1 - b^2 \delta^2). \label{eq:sn-variance}
  \end{align*}
}
\frame{
  \frametitle{The Skew-Normal Distribution}

  The $SN(0,1,\lambda)$ distribution is called the standard skew-normal; its pdf
  is

  \begin{equation} \label{eq:standard-sn-pdf}
    f_Z(x|\lambda) = 2 \cdot \phi(x) \cdot \Phi (\lambda x), \quad x \in \R.
  \end{equation}

  Similar to the normal and standard normal, $Z = \frac{Y - \mu}{\sigma}$ and $Y
  = \sigma Z + \mu$.
}
\frame{
  \frametitle{The Skew-Normal Distribution}

  \begin{align*}
    f(x|\mu, \sigma, \lambda=0) &= \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot \Phi(0) \\
    &= \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot 0.5 \\
    &= \frac1\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \\
    &= \frac1\sigma \cdot \frac{1}{\sqrt{2\pi}} \;\cdot\; \exp \left( -\frac{(x-\mu)^2}{2\sigma^2} \right) \\
    &= \frac{1}{\sqrt{2\pi}\sigma} \;\cdot\; \exp \left( -\frac{(x-\mu)^2}{2\sigma^2} \right),
  \end{align*}

  which is the pdf of the normal distribution ($\mu$, $\sigma$). Furthermore,
  when $\lambda > 0$, the curve skews to the right, and when $\lambda < 0$, it
  skews to the left (a property we will prove in section
  \ref{subsec:four-properties}).
}
\frame{
  \frametitle{The Skew-Normal Distribution}

  \begin{property} \label{prop:1}
    If $Z \sim SN(0, 1, \lambda)$, then $(-Z) \sim SN(0, 1, -\lambda)$.
  \end{property}

  \begin{property} \label{prop:2}
    If $Z \sim SN(0, 1, \lambda)$, then $Z^2 \sim \chi^2_1$ (chi-square with 1 degree of freedom).
  \end{property}

  \begin{property} \label{prop:3}
    As $\lambda \to \pm \infty$, \thinspace $SN(0,1,\lambda)$ tends to the half normal distribution, $\pm |N(0,1)|$.
  \end{property}

  \begin{property} \label{prop:4}
    The moment generating function of $SN(0,1,\lambda)$ is

    \begin{equation} \label{eq:p4-sn-mgf}
      M(t|\lambda) = 2 \cdot \Phi (\delta t) \cdot e^{t^2/2},
    \end{equation}

    where $\delta = \frac{\lambda}{\sqrt{1 + \lambda^2}}$ and $t \in (-\infty, \infty)$.
  \end{property}
}
\frame{
  \frametitle{The Skew-Normal Distribution}
}