### 2011-02-14 project/presentation/skew-normal.tex

% The Skew-Normal Distribution
\frame{ \frametitle{The Skew-Normal Distribution: Foundations}
  \begin{definition}[Skew-normal]
    Let $Y$ be a skew-normal distribution, with location parameter $\mu \in
    \R$, scale parameter $\sigma > 0$, and shape parameter $\lambda \in \R$.
    Then $Y$ has pdf

    \begin{equation*}
      f(x|\mu, \sigma, \lambda) = \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot \Phi \left( \frac{\lambda(x-\mu)}{\sigma} \right), \quad x \in \R,
    \end{equation*}

    where $\phi$ is the standard normal pdf and $\Phi$ is the standard normal
    cdf.

    \vsep
    We write $Y \sim SN(\mu, \sigma, \lambda)$.
  \end{definition}
}
\frame{ \frametitle{The Skew-Normal Distribution: Foundations}
  \begin{lem*}
    If $f_0$ is a one-dimensional probability density function symmetric about
    0, and $G$ is a one-dimensional distribution function such that $G'$ exists
    and is a density symmetric about 0, then

    \begin{equation*}
      f(z) = 2 \cdot f_0(z) \cdot G\{w(z)\} \quad (-\infty < z < \infty)
    \end{equation*}

    is a density function for any odd function $w(\cdot)$. (Lemma 1,
    \citealp{azzalini})
  \end{lem*}
}
\frame{ \frametitle{The Skew-Normal Distribution: Foundations}

  Basic properties:

  \begin{align*}
    E(Y) &= \mu + b \delta \sigma \\
    E(Y^2) &= \mu^2 + 2b \delta \mu \sigma + \sigma^2 \\
    E(Y^3) &= \mu^3 + 3 b \delta \mu^2 \sigma + 3 \mu \sigma^2 + 3 b \delta \sigma^3 - b \delta^3 \sigma^3 \\
    Var(Y) &= \sigma^2 (1 - b^2 \delta^2)
  \end{align*}

  where $b = \sqrt{\frac{2}{\pi}}$ and $\delta = \frac{\lambda}{\sqrt{1 +
  \lambda^2}}$. \citep{pewsey}
}
\frame{ \frametitle{The Skew-Normal Distribution: Foundations}

  What happens when $\lambda=0$?

  \begin{align*}
    f(x|\mu, \sigma, \lambda=0) &= \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot \Phi(0) \\
    &= \frac2\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \cdot 0.5 \\
    &= \frac1\sigma \cdot \phi \left( \frac{x-\mu}{\sigma} \right) \\
    &= \frac{1}{\sqrt{2\pi}\sigma} \;\cdot\; \exp \left( -\frac{(x-\mu)^2}{2\sigma^2} \right),
  \end{align*}

  which is the pdf of the normal distribution ($\mu$, $\sigma$).
}
\frame{ \frametitle{The Skew-Normal Distribution: The Standard Skew-Normal}

  \begin{definition}[Standard skew-normal]
    The $SN(0,1,\lambda)$ distribution is called the standard skew-normal and
    has pdf

    \begin{equation*}
      f_Z(x|\lambda) = 2 \cdot \phi(x) \cdot \Phi (\lambda x), \quad x \in \R.
    \end{equation*}
  \end{definition}

  Similar to the normal and standard normal, $Z = \frac{Y - \mu}{\sigma}$ and $Y
  = \sigma Z + \mu$.
}
\frame{ \frametitle{The Skew-Normal Distribution: The Standard Skew-Normal}

  \begin{property}
    If $Z \sim SN(0, 1, \lambda)$, then $(-Z) \sim SN(0, 1, -\lambda)$.
  \end{property}
  \vsep
  \begin{property}
    If $Z \sim SN(0, 1, \lambda)$, then $Z^2 \sim \chi^2_1$ (chi-square with 1 degree of freedom).
  \end{property}
}
\frame{ \frametitle{The Skew-Normal Distribution}
  \frametitle{The Skew-Normal Distribution: The Standard Skew-Normal}

  \begin{property}
    As $\lambda \to \pm \infty$, \thinspace $SN(0,1,\lambda)$ tends to the half normal distribution, $\pm |N(0,1)|$.
  \end{property}
  \vsep
  \begin{property}
    The moment generating function of $SN(0,1,\lambda)$ is

    \begin{equation*}
      M(t|\lambda) = 2 \cdot \Phi (\delta t) \cdot e^{t^2/2},
    \end{equation*}

    where $\delta = \frac{\lambda}{\sqrt{1 + \lambda^2}}$ and $t \in (-\infty, \infty)$.
  \end{property}
}